---
title: "Lime Adoption Summary"
author: "Matt Lowes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Quickly summarize lime adoption so far in Nambale only

```{r import}
wd <- "/Users/mlowes/drive/soil health study/data/ke baseline"
dd <- paste(wd, "data", sep = "/")
od <- paste(wd, 'output', sep = "/")

d <- read.csv(paste(dd, "KE SHS Baseline Survey Data.csv", sep = '/'))
soil <- read.csv(paste(dd, "Kenya_SHS_results.csv", sep = "/"))
Identifiers <- read.csv(paste(dd, "Combined meta with SSN.csv", sep = '/'), stringsAsFactors = F)
```

First step is merging the data. The ids in the Identifiers data need to be adjusted so that they match the ids in the ids in CommCare

```{r}
Identifiers$oaf.ID <- tolower(Identifiers$oaf.ID)
Identifiers$DISTRICT[Identifiers$DISTRICT=="Kakamega North"] <- "Kakamega B (North)"
Identifiers$DISTRICT[Identifiers$DISTRICT=="Kakamega South"] <- "Kakamega (South)"


d$id.Soil_Sample_Id <- tolower(d$id.Soil_Sample_Id)

Identifiers$sample_id <- ifelse(grepl("c", Identifiers$oaf.ID), Identifiers$oaf.ID, paste(tolower(Identifiers$DISTRICT), Identifiers$oaf.ID, sep = ""))

Identifiers$sample_id <- gsub(" ", "", Identifiers$sample_id)
d$id.Soil_Sample_Id <- gsub(" ", "", d$id.Soil_Sample_Id)
```

```{r}
table(d$id.Soil_Sample_Id %in% Identifiers$sample_id)
```

We're currently missing `r table(d$id.Soil_Sample_Id %in% Identifiers$sample_id)[1]` matches. Investigate why we're missing these. Below are the ids that do not appear in the Identifiers data but appear in the CommCare survey data:

```{r}
d$id.Soil_Sample_Id[!d$id.Soil_Sample_Id %in% Identifiers$sample_id]
```
Issues seem to be:

* KKM south - resolved, see block above
* KKM north - resolved, see block above
* assorted unmatched values from other districts

These are the ids that appear in the Identifiers data but not in the CommCare surey data:




```{r}
Identifiers$sample_id[!Identifiers$sample_id %in%  d$id.Soil_Sample_Id]
```

```{r}
d <- merge(d, Identifiers[,c("SSN", "sample_id")], by.x="id.Soil_Sample_Id", by.y="sample_id", all.x=TRUE)
```






