---
title: "Rwanda Soil Health Study - Round 2 Check"
author: "[Matt Lowes](mailto:matt.lowes@oneacrefund.org)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_notebook:
    number_sections: yes
    code_folding: show
    fig_caption: yes
    fig_height: 6
    theme: flatly
    toc: yes
    toc_depth: 6
    toc_float: yes
subtitle: "On-going data check"
---
```{r echo=F, message=FALSE}
#### set up
## clear environment and console
rm(list = ls())
cat("\014")

## set up some global options
# always set stringsAsFactors = F when loading data
options(stringsAsFactors=FALSE)

# show the code
knitr::opts_chunk$set(echo = TRUE)

# define all knitr tables to be html format
options(knitr.table.format = 'html')

# change code chunk default to not show warnings or messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

## load libraries
# dplyr and tibble are for working with tables
# reshape is for easy table transformation
# knitr is to make pretty tables at the end
# ggplot2 is for making graphs
# readxl is for reading in Excel files
# MASS is for running boxcox tests
# gridExtra is for arranging plots
# cowplot is for adding subtitles to plots
# robustbase is to run robust regressions to compensate for outliers
# car is for performing logit transformations
libs <- c("dplyr", "reshape2", "knitr", "ggplot2", "tibble", "readxl", 
    "MASS", "gridExtra", "cowplot", "robustbase", "car", "knitr", "DT")
lapply(libs, require, character.only = T, quietly = T, warn.conflicts = F)

#### define helpful functions
# define function to adjust table widths
html_table_width <- function(kable_output, width) {
  width_html <- paste0(paste0('<col width="', width, '">'), collapse = "\n")
  sub("<table>", paste0("<table>\n", width_html), kable_output)
}
```

# Objective

Perform basic data checks on Rwanda round soil health study data to provide real time feedback to Nathaniel and enumeration teams to maximize the value of the teams in the field.

# Key takeaways

> See [categorical graphs](#categorical-var-graphs)

> See [numerical variable graphs](#numeric-var-graphs)

> See [variables for follow up](#specific-points)

# Data

## Baseline and round 1

```{r}
load("../rw_round_1/fieldDat_final.Rdata")
```

## Round 2 data
Load data and make variable names nice. `ytwo` is the complete data. `yt` is the data I'll be manipulating.
```{r loading data, message=FALSE}
dataDir <- normalizePath(file.path("..", "..", "data"))
source("../oaflib/commcareExport.R")
source("../oaflib/misc.R")
#source("../oaflib/ccSluthing.R")

ytwo <- getFormData("oafrwanda", "M&E", "17B Ubutaka (Soil)", forceUpdate=F)

names(ytwo) <- tolower(make.names(names(ytwo)))

yt <- ytwo %>%
  setNames(gsub("x.form.", "", names(.))) %>%
  mutate(sample_id = tolower(sample_id))
```

# Observation count

## Missing data
Reality check some observations. Are they completely missing? 
```{r}
blankRow <- apply(yt, 1, function(x){
  sum(x=="" | is.na(x))
})

naRow <- data.frame(table(round(blankRow/dim(yt)[2],2)*100)) %>%
  mutate(Var1 = as.numeric(as.character(Var1))) %>%
  filter(Var1>=25) %>%
  mutate(Var1 = paste0(Var1, "%"))
kable(naRow, col.names = c("% of columns missing values", "Freq"),format="markdown")
```

Some observations have `r as.character(naRow$Var1[nrow(naRow)])` of observations missing. What's going on with these? I'm only showing rows with greater than 25% missing. It's somewhat arbitrary but to not have data for a fourth of the survey seems odd.

## Matches with previous rounds

```{r}
#table(yt$sample_id %in% unique(fieldDat$sample_id))
```

> Through `r format(Sys.time(), '%B %d, %Y')` we have one soil sample id in round 2 that doesn't exist in previous rounds. It is sample_id `r yt$sample_id[!yt$sample_id %in% unique(fieldDat$sample_id)]`.

What is up with that observation?

## Completion rates

```{r}
baseComplete <- fieldDat %>%
  filter(season =="15b") %>%
  dplyr::select(sample_id) 

compB <- round(prop.table(table(baseComplete$sample_id %in% yt$sample_id)),2)

r1Complete <- fieldDat %>%
  filter(season =="16b") %>%
  dplyr::select(sample_id) 


compR1 <- round(prop.table(table(r1Complete$sample_id %in% yt$sample_id)),2)
```

> Through `r format(Sys.time(), '%B %d, %Y')` we've found `r compB[[2]]*100`% of baseline farmers. We have `r nrow(baseComplete)-table(baseComplete$sample_id %in% yt$sample_id)[[2]]` to go.

> Through `r format(Sys.time(), '%B %d, %Y')` we've found `r compR1[[2]]*100`% of round 1 farmers. We have `r nrow(r1Complete)-table(r1Complete$sample_id %in% yt$sample_id)[[2]]` to go.

As we get closer to the end of enumeration I will produce a report of sample_ids that have not yet been surveyed so we can understand precisely why some farmers were not found again.

## Attrition summary

```{r}
# code to check merge match before match is performed
# the code should show you how many overlaps you have in your merge variable
# and a summary of the values that don't match

# tested using ke_round_1_py

vecFiller <- function(vec){
  vec[2] = ifelse(is.na(vec[2]), 0, vec[2])
  return(vec)
}


mergeReport <- function(master, using){
  
  # dat1 into dat2
  
  c1 <- vecFiller(table(master %in% using))
  c2 <- vecFiller(table(using %in% master))
  
  notMatch1 <- master[!master %in% using]
  notMatch2 <- using[!using %in% master]
  
  tab <- data.frame(rbind(c1, c2))
  rownames(tab) <- c("Master in using", "Using in Master")
  
  outList <- list(tab, notMatch1, notMatch2)
  names(outList) <- c("matching.table", "missing.from.using", "missing.from.master")
  
  return(
    outList
  )
  
  
}

attReport <- mergeReport(baseComplete$sample_id, yt$sample_id)
```

We currently have `r length(attReport$missing.from.using)` more surveys to complete. The complete list of missing farmer ids will be shared once we're down to a more reasonable number.


# Data reality check

The variable names are not great. I think it's due to the CommCare API export. I'm going to see if there's an easier way to have uesable variable names. I've creaetd a CC export titled "M&E - Soil Health Study - 17B Ubutaka (Soil)" on August 8th. I'm only using it to get more usable variable names quickly. Ideally there'd be away to get that info directly from CommCare.

```{r}
varNames <- read_csv("M&E - Soil Health Study - 17B Ubutaka (Soil) (2017-08-04) 2017-08-04.csv") %>% 
  setNames(gsub("form.", "", names(.))) %>%
  setNames(gsub("farmer.", "", names(.))) %>%
  setNames(gsub("prim_17a_.", "", names(.))) %>%
  setNames(gsub("sec_17a_.", "", names(.))) %>%
  setNames(gsub("prim_17b_.", "", names(.))) %>%
  setNames(gsub("sec_17b_.", "", names(.))) %>%
  setNames(gsub("ammend_17a_.", "", names(.))) %>%
  setNames(gsub("ammend_17b_.", "", names(.))) %>% 
  as.data.frame()


names(yt)[11:112] <- names(varNames)[3:104]
# rename one variable to remove duplicates
names(yt)[31] <- "nameOne"
names(yt) <- tolower(names(yt))
# manually fix a couple other names
names(yt)[53] <- "crop1_yield_comparison_17a"
names(yt)[58] <- "crop2_yield_comparison_17a"

```

## Categorical variables

```{r}
catVars <- names(yt)[sapply(yt, function(x){
  is.character(x)
})]

enumClean <- function(dat, x, toRemove){
  dat[,x] <- ifelse(dat[,x] %in% toRemove, NA, dat[,x])
  return(dat[,x])
}

strTable <- function(dat, x){
  varName = x
  tab = as.data.frame(table(dat[,x], useNA = 'ifany'))
  tab = tab[order(tab$Freq, decreasing = T),]
  end = ifelse(length(tab$Var1)<10, length(tab$Var1), 10)
  repOrder = paste(tab$Var1[1:end], collapse=", ")
  out = data.frame(variable = varName,
                   responses = repOrder)
  
  return(out)
}

# clean up known values
catEnumVals <- c("-99", "-88", "- 99", "-99.0", "88", "_88", "- 88", "0.88",
                 "--88", "__88", "-88.0", "99.0")
yt[,catVars] <- sapply(catVars, function(y){
  yt[,y] <- enumClean(yt,y, catEnumVals)
})

# put all categorical values in lower
yt[,catVars] <- sapply(yt[,catVars], function(x) tolower(x))

responseTable <- do.call(rbind, lapply(catVars, function(x){
  strTable(yt, x)
}))
```

### Table of categorical responses

```{r}
kable(responseTable, format="markdown")
```

### Categorical var graphs

* district_17b names are spelled wrong. How is that possible? Do we care?

```{r}
repGraphs <- function(dat, x){
  tab = as.data.frame(table(dat[,x], useNA = 'ifany'))
  tab = tab[order(tab$Freq, decreasing = T),]
  print(
    ggplot(data=tab, aes(x=Var1, y=Freq)) + geom_bar(stat="identity") +
      theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
      labs(title =paste0("Composition of variable: ", x))
  )
}

adminVars <- tolower(c(names(yt)[grep("meta", names(yt))],"enum_name", "text_final.photo",  "participation", "refusal", "phone",  "comment", "gps", "sample_id", "sampling.barcode", "id", "domain", "date_header", "form.case..case_id", "site", "district1", "site1", "plot.location", "New_soil_sample_id", "#form/Sampling2017_Complete", "appformid",names(yt)[grep("soil_id", names(yt))], "nameOne", "tel", "tel_respondent", "neighbor_tel", "finish_time", "other_comments", "Description", "neighbor_or_voisin_phone", "name", "cell_field_17b", "name_respondent","respondent_in_16b","gps_hidden", "farmer_phone", "respondent_phone", "village", "cell", names(yt)[grep("enumerator", names(yt))], "start_time", "village_17b", "cell_input_17b", "not_find_why_other", "compost_other_17a"))
nonAdminVars <- catVars[!catVars %in% adminVars]

for(i in 1:length(nonAdminVars)){
  repGraphs(yt, nonAdminVars[i])
}
```
## Numeric variables

```{r}
numVars <- names(yt)[sapply(yt, function(x){
  is.numeric(x)
})]
```

Basic cleaning of known issues like enumerator codes for DK, NWR, etc.
```{r}
enumVals <- c(-88,-85, -99, -9)

yt[,numVars] <- sapply(numVars, function(y){
  yt[,y] <- enumClean(yt,y, enumVals)
  yt[,y] <- as.numeric(yt[,y])
})
```

### Numeric var graphs

```{r}
temp <- numVars


noNeed <- c("intro.intro", "text_final.d_photo", "text_final.d_soilsample")
numVars <- numVars[!numVars %in% noNeed]
# remove variables that are entirely missing
totallyMissing <- sapply(yt[,numVars], function(x){
  all(is.na(x))
  })

numVars <- numVars[totallyMissing==F]


for(i in 1:length(numVars)){
    if(length(unique(yt[,numVars[i]]))>10){
      print(ggplot(data=yt, aes(x=yt[,numVars[i]])) +  
              geom_density() + 
              theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) + 
              labs(x = numVars[i])
            )
    } else{
    print(ggplot(data=yt, aes(x=yt[,numVars[i]])) + 
            geom_histogram(stat="count") + 
            theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
            labs(x = numVars[i])
      )
    }
    #multiplot(temp1, temp2, cols = 2
}
```

# Follow up

```{r message=F}
detectOutlier <- function(dat, x){

# function checks distribution and if lognormal, coverts and outputs
# outliers from log normal distribution. have to remove 0s to check
  fits <- list(
 no = fitdistr(dat[,x][complete.cases(dat[,x])],"normal"),
 lo = fitdistr((dat[,x])[complete.cases(dat[,x]) & dat[,x]!=0 & dat[,x]>0],"log-normal")
 #ca = fitdistr(dat[,x][complete.cases(dat[,x])],"cauchy"),
 #we = fitdistr(dat[,x][complete.cases(dat[,x])], "weibull")
 )
# get the logliks for each model...
fitCheck <- sapply(fits, function(i) i$loglik)
bestFit <-names(which.max(fitCheck))

trans <- sapply(dat[,x], function(y){
  ifelse(bestFit=="lo", log10(y+1), y)
})
  
  
  q1 = summary(trans)[[2]]
  q3 = summary(trans)[[5]] 
  iqr = q3-q1
  mark  = ifelse(trans < (q1 - (1.5*iqr)) | trans > (q3 + (1.5*iqr)), 1,0)
  # tab = rbind(
  #   summary(d[,x]),
  #   summary(d[mark==0, x])
  # )
  out = dat[mark==1, c("district", "cell", "enumerator", "sample_id", "farmer_phone", x)]
  out = melt(out, measure.vars = x)
  out = out[!is.na(out$district),]
  
  return(out)

}

printIQR <- do.call(rbind, lapply(numVars, function(y){
  #print(y)
  return(detectOutlier(yt, y))
}))

printIQR <- printIQR[order(printIQR$enumerator,printIQR$cell, printIQR$sample_id),]
```

## Specific points

The initial output might still be showing values that the M&E team feels are reasonable values in reality despite being flagged as statistical outliers. 

I've included a couple examples of how we can filter the data through the code to arrive at a well documented list of values to investigate.

* `compost_17a` many farmers list 0. It's plausible but perhaps worth a general follow up?
Although the key here should be if farmers are not applying, it should be NA. If they are applying, are they really only applying small quantities? Look into this.
* `kg_seed_crop2_17a` appears so often because I think most farmers said 0 because they have only one crop. If the values are sensible then we can let it go. With a bit more work I could look at only non-zero values.
* `field_area_17b` has some large values. I think these were misentered. Clarify with the enumerator and we can just fix them here.
* Large kg_seed values? These are probably better reality checked as kg/are. I can convert to field sizes
* 22 goats?
* yield variable seems quite high as well. What constraints are currently in place? Do we have enumerators falling outside them? Can we tighten them down?

```{r}
modifyOutlierOutput <- function(dat, varname, vals){
  dat = dplyr::filter(dat, variable!=varname & (!value %in% vals))
  return(dat)
}

printIQR <- modifyOutlierOutput(printIQR, "found", 1)
printIQR <- modifyOutlierOutput(printIQR, "compost", 1)
printIQR <- modifyOutlierOutput(printIQR, "compost_17a", 0)
printIQR <- modifyOutlierOutput(printIQR, "n_chickens", 2:7)


```

**Read the values in this table with the above histograms and density plots in mind. Not all the values listed in the table will be impossible values but they were flagged for being far from the central mass of values we have on that data point**

```{r}
datatable(printIQR)
```

```{r}
write.csv(printIQR, file=paste0("output/", "rw_shs_r2 vars to check update",format(Sys.time(), '%B %d %Y'),  ".csv"), row.names = F)

for(i in 1:length(unique(printIQR$enumerator))){
  outSheet <- printIQR[printIQR$enumerator==unique(printIQR$enumerator)[i], ]
  write.csv(outSheet, file=paste("output",
          paste("rw_shs_r2 vars to check ", unique(printIQR$enumerator)[i], ".csv", sep = ""), sep = "/"))
  
}
```
