---
title: "Kenya Soil Health Study Analysis"
author: '[Matt Lowes](mailto:matt.lowes@oneacrefund.org)'
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_notebook:
    number_sections: yes
    code_folding: show
    theme: flatly
    toc: yes
    toc_depth: 6
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
rm(list = ls())
cat("\014")

## set up some global options
# always set stringsAsFactors = F when loading data
options(stringsAsFactors=FALSE)

# show the code
knitr::opts_chunk$set(echo = TRUE)

# define all knitr tables to be html format
options(knitr.table.format = 'html')

# change code chunk default to not show warnings or messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

libs <- c("dplyr", "reshape2", "knitr", "ggplot2", "tibble", "readxl", 
    "MASS", "gridExtra", "cowplot", "robustbase", "car", "RStata", "foreign",
    "tidyr", "readxl", "stringr", "sp", "dismo", "leaflet", "XML", "ggmap", "knitr",
    "MASS")
lapply(libs, require, character.only = T, quietly = T, warn.conflicts = F)

#### define helpful functions
# define function to adjust table widths
html_table_width <- function(kable_output, width) {
  width_html <- paste0(paste0('<col width="', width, '">'), collapse = "\n")
  sub("<table>", paste0("<table>\n", width_html), kable_output)
}

source("../oaflib/commcareExport.R")
source("../oaflib/misc.R")
```

# Background

This file will pick up on from the Kenya round 1 file and finish the analysis. The cleaning of the data up through the round 1 data (baseline and round 1) happens in the `kenya_round_1` folder. Going forward, the cleaning of new data will happen in the `shs_cleaning.R` file, which I'll source here, and then the analysis will happen here.

# Objective

Analyze the soil health study data.

# Data
```{r}
# load the latest kenya data from the cleaning file >> or maybe do the combining there.
load()


```


# Output

# Appendix
